Place of death in NRS COVID-19 deaths

### 2021-08-16 covid_nrs_place_of_death_data.Rmd
added proportion of yearly deaths broken down by geography & started plotting visualisation;
also added some spaces to the save_output_file function to make it easier to read

>>>>

## By geography

```{r}
proportions_of_deaths_by_place_by_geography <-
  bind_rows(
    ## historic deaths
    bind_rows(  # merge HB & LA past deaths, and tag them appropriately
      past_deaths_hb %>% mutate(geography = "hb"),
      past_deaths_la %>% mutate(geography = "la")
    ) %>%
    group_by(year, place_of_death, ref_area, geography) %>% summarise(number_of_deaths = sum(number_of_deaths), .groups="drop") %>%  # compute the total number of yearly deaths by place of death & geography
    filter(place_of_death != "All" & !is.na(as.integer(year))) %>%  # remove blank/irrelevant entries
    group_by(year, ref_area, geography) %>%
    mutate(
      total_deaths = sum(number_of_deaths)  # compute total annual deaths across place of death, but within each geographical area & year
    ) %>%
    ungroup %>%
    mutate(proportion_of_total = number_of_deaths / total_deaths, year=as.integer(year))  # compute proportion of deaths in any palce relative to total annual deaths in geographical area
    ,
    ## current deaths
    bind_rows(  # merge HB & LA past deaths, and tag them appropriately
      total_deaths_hb %>% mutate(geography = "hb"),
      total_deaths_la %>% mutate(geography = "la")
    ) %>%
    select(year, place_of_death, number_of_deaths = deaths_all_causes, geography, ref_area) %>%  # keep only relevant columns & rename deaths due to all causes
    group_by(year, place_of_death, ref_area, geography) %>% summarise(number_of_deaths = sum(number_of_deaths), .groups="drop") %>%  # compute the total number of yearly deaths by place of death & geography
    filter(place_of_death != "All" & !is.na(as.integer(year))) %>%  # remove blank/irrelevant entries
    group_by(year, ref_area, geography) %>%
    mutate(
      total_deaths = sum(number_of_deaths)  # compute total annual deaths across place of death, but within each geographical area & year
    ) %>%
    ungroup %>%
    mutate(proportion_of_total = number_of_deaths / total_deaths, year=as.integer(year))  # compute proportion of deaths in any palce relative to total annual deaths in geographical area
  )
```

### 2021-08-16 covid_nrs_place_of_death_visualisations.Rmd
added proportion of yearly deaths broken down by geography & started plotting visualisation;
also added some spaces to the save_output_file function to make it easier to read

library(tidyverse)  # for tidy workflow
library(opendatascot)  # importing data from ScotGov open data website
library(phsmethods)  # methods for working with PHS data
library(readxl)  # for reading excel files
# library(SPARQL)  # taken care of by opendatascot
library(lubridate)  # dealing with dates
library(janitor)  # for cleaning column names
library(ISOweek)  # for computing date from ISO week number + year
library(sf)  # for mapping
library(ggrepel)  # for 'mapping'repelling' labels/texst in ggplot
library(patchwork)  # for assembling plots
library(extrafont)  # for working with fonts
library(openxlsx)  # for creating xlsx files
library(knitr)  # for displaying tables in rmd
```

@@

```{r}
figure_proportion_home_deaths_over_time <-
  merged_proportions_of_deaths_by_place %>%
  mutate(place_of_death = factor(x = place_of_death, levels = order_of_place_of_death_levels[c(1,3,2,4)])) %>%  # move home to 2nd position to match order in plot
  ggplot(., aes(x=year, y=proportion_of_total, colour=place_of_death)) +
  geom_point() +
  geom_line(size = 1) +
  geom_text(data = ~filter(.x, year==2021), aes(x=2021.1, y=proportion_of_total, label="*"), inherit.aes = FALSE) +
  scale_x_continuous(breaks = 2015:2021) +
  scale_y_continuous(label = scales::percent) +
  scale_colour_viridis_d(option="C") +
  # theme_minimal() +
  theme(legend.position = "top", plot.caption = element_text(size = 10, colour = "gray60")) +
  labs(
    x = "Year", y= "Proportion of yearly deaths (%)",
    colour = "",
    subtitle = paste0("* ",only_2021_text),
    caption = paste0(c(source_nrs_text),collapse="\n")
    ) +
  guides(colour = guide_legend(nrow=2)) +  # spread colour legend over two rows
  NULL

save_output_file(filename = paste0(dir_outputs,"/figure_proportion_home_deaths_over_time"), extensions = c(".pdf",".svg",".png"), plot = figure_proportion_home_deaths_over_time, width = 12, height = 10, units = "cm", dpi = 300)

figure_proportion_home_deaths_over_time
```

@@

>>>>

library(tidyverse)  # for tidy workflow
library(opendatascot)  # importing data from ScotGov open data website
library(phsmethods)  # methods for working with PHS data
library(readxl)  # for reading excel files
# library(SPARQL)  # taken care of by opendatascot
library(lubridate)  # dealing with dates
library(janitor)  # for cleaning column names
library(ISOweek)  # for computing date from ISO week number + year
library(sf)  # for mapping
library(ggrepel)  # for 'mapping'repelling' labels/texst in ggplot
library(patchwork)  # for assembling plots
library(extrafont)  # for working with fonts
library(openxlsx)  # for creating xlsx files
library(knitr)  # for displaying tables in rmd
library(lemon)  # for showing axis & axis labels in every facet in ggplot
```

@@

```{r}
(figure_proportion_home_deaths_over_time <-
  merged_proportions_of_deaths_by_place %>%
  mutate(place_of_death = factor(x = place_of_death, levels = order_of_place_of_death_levels[c(1,3,2,4)])) %>%  # move home to 2nd position to match order in plot
  ggplot(., aes(x=year, y=proportion_of_total, colour=place_of_death)) +
  geom_point() +
  geom_line(size = 1) +
  geom_text(data = ~filter(.x, year==2021), aes(x=2021.1, y=proportion_of_total, label="*"), inherit.aes = FALSE) +
  scale_x_continuous(breaks = 2015:2021) +
  scale_y_continuous(label = scales::percent) +
  scale_colour_viridis_d(option="C") +
  # theme_minimal() +
  theme(legend.position = "top", plot.caption = element_text(size = 10, colour = "gray60")) +
  labs(
    x = "Year", y= "Proportion of yearly deaths (%)",
    colour = "",
    subtitle = paste0("* ",only_2021_text),
    caption = paste0(c(source_nrs_text),collapse="\n")
    ) +
  guides(colour = guide_legend(nrow=2)) +  # spread colour legend over two rows
  NULL
)

save_output_file(filename = paste0(dir_outputs,"/figure_proportion_home_deaths_over_time"), extensions = c(".pdf",".svg",".png"), plot = figure_proportion_home_deaths_over_time, width = 12, height = 10, units = "cm", dpi = 300)
```

@@

# Proportions of home deaths by geography

## HB

```{r}
# (figure_proportion_home_deaths_over_time <-
  proportions_of_deaths_by_place_by_geography %>%
  filter(geography == "hb") %>%
  mutate(place_of_death = factor(x = place_of_death, levels = order_of_place_of_death_levels[c(1,3,2,4)])) %>%  # move home to 2nd position to match order in plot
  ggplot(., aes(x=year, y=proportion_of_total, colour=place_of_death)) +
  geom_point() +
  geom_line(size = 1) +
  geom_text(data = ~filter(.x, year==2021), aes(x=2021.1, y=proportion_of_total, label="*"), inherit.aes = FALSE) +
  # facet_wrap(~ ref_area, scales = "free_x") +
  facet_rep_wrap(~ ref_area, repeat.tick.labels = TRUE) +
  scale_x_continuous(breaks = 2015:2021) +
  # scale_y_continuous(breaks = seq(0.1, 0.6, by = 0.1), label = function(x) scales::percent(x, accuracy = 1)) +
  scale_y_continuous(label = function(x) scales::percent(x, accuracy = 1)) +
  scale_colour_viridis_d(option="C") +
  # theme_minimal() +
  theme(
    legend.position = "top", 
    plot.caption = element_text(size = 10, colour = "gray60"),
    panel.border = element_rect(fill = NA, colour = "gray80"),  # draw a transparent rectangle as the border
    strip.background = element_rect(fill = NA, colour = "gray80")
    ) +
  labs(
    x = "Year", y= "Proportion of yearly deaths (%)",
    colour = "",
    subtitle = paste0("* ",only_2021_text),
    caption = paste0(c(source_nrs_text),collapse="\n")
    ) +
  guides(colour = guide_legend(nrow=2)) +  # spread colour legend over two rows
  NULL
# )

save_output_file(filename = paste0(dir_outputs,"/figure_proportion_home_deaths_over_time"), extensions = c(".pdf",".svg",".png"), plot = figure_proportion_home_deaths_over_time, width = 12, height = 10, units = "cm", dpi = 300)
```
